\sys: Automated Data-Parallel Shell Processing

### 2020-05-03 pldi/draft.tex
Minor updates in eval

\subsection{Classic \unix One-liners}
\label{ours}

@@

\heading{Programs}
In a recent event celebrating \unix's 50-year legacy, Bell Labs created a set of challenges~\cite{unix50}.

solving challenges by

ability to compose larger programs in versatile ways from smaller utility programs
summarized by McIllroy's Unix philosophy memo.

The ability to compose larger programs in versatile ways from smaller utility programs was a hallmark of the \unix design, and the game's levels are designed to stress that.

This section reports on programs that were run unaltered---highlighting 
These are pipelines that were run verbatim
In cases where an obvious fix improves the pipeline's performance, we report on both \sys's original speedup and the one achieved by applying the fix.

\todo{Motivate why unix50 is a nice
  set of pipelines.} We took the solutions for the unix50 from
\todo{Ref} and we executed them \emph{without any modification} with

@@

the problems are well designed to highlight the \unix philosophy.
pipelines were found out in the wild and are not necessarily written in a particularly

@@

\heading{\ttt{sort --parallel}}
We compare a single \todo{GNU} \ttt{sort} optimized by \sys with the
same \ttt{sort} with the \ttt{--parallel} flag set. The results are
shown in \Cref{fig:sort-parallel-comparison}. There are a couple
important points to note. First of all, \sys without the \ttt{eager}
optimization enabled performs comparably with \ttt{sort
  --parallel}. On the other hand, enabling the \ttt{eager}
optimization allows \sys to perform much better (almost \todo{2}
times) than \ttt{sort} with the \ttt{--parallel} flag. That is because
\sys with the \ttt{eager} optimization enabled essentially uses
intermediate buffers (on disk) between the merge phases. On the other
hand, \ttt{sort --parallel} sorts the input trying to minimize
additional memory (so doesn't allocate intermediate buffers). \kk{If
  we want, we can add here numbers where we allow sort --parallel to
  use additional ram for intermediate buffers (using the -S flag).} Of
course, it would be possible to manually create a longer pipeline and
use intermediate buffers to speed this up, but this defeats the
purpose of having a simple solution.

@@

\heading{GNU Parallel}
We compare \sys to \ttt{parallel} (v.20160422), a GNU utility for running other commands in parallel~\cite{Tange2011a}.
In the first experiment, we leverage \ttt{parallel}'s \ttt{--jobs} flag and redirect input from a file to 
Sequential execution takes \todo{XX}s, and \sys's parallel execution takes \todo{XX}s (speedup: \todo{XX}\%).
We note that this pipeline is harsh for \sys, in that most of the overhead comes from a single command, masking \sys's improvements in other parts of the pipeline.

@@

There are a few possible ways one might attempt to use GNU \ttt{parallel} to parallelize this program, which compared to \sys lead to a combination of smaller performance improvements, incorrect execution with respect to the sequential program, and higher rewriting effort.
Assuming a user knows that the bottleneck comes from the \ttt{XX} stage, they could attempt to use \ttt{parallel} there leading to a runtime of \todo{XX}s (speedup: \todo{XX}\%)---worse than surprisingly worse than.

Alternatively, a user could (incorrectly) sprinkle \ttt{parallel} across the entire program.
This is a common strategy as GNU \ttt{parallel} allows feeding an entire script for parallelization via \ttt{stdin}. 
Doing this leads to \todo{XXX}\% performance improvements but severely incorrect results with respect to the sequential execution---a \ttt{diff} that averages about \todo{XX--XX} lines.
\sys's conservative program transformations will not attempt parallelizing program sections that contain commands with unclear \todo{parallelizability} properties.
	% or commands whose inputs depend on earlier outputs.
GNU \ttt{parallel} is oblivious to command semantics, risking breaking program semantics as in this case.

\heading{Manual Parallelization}
In an attempt to highlight how a user might attempt to manually parallelize individual commands using the abstractions provided by the POSIX shell.
The benefits depend on several factors, such as whether a command is I/O- or CPU-bound or whether but since

Merging is provided as Python programs

@@

\heading{Take-away}
\sys will choose not to improve performance rather than breaking correctness; as a result, it does not always improve performance.
In highly-parallel benchmarks, it can lead to performance that is comparable to highly specialized streaming frameworks---at a fraction of the cost in terms of manual development.

>>>>

\subsection{Common \unix One-liners}
\label{ours}

The goal of this subsection is to evaluate \sys on popular, common, or classic \unix pipeline patterns.

@@

\heading{Programs}
In a recent event celebrating \unix's 50-year legacy, Bell Labs created a set of \todo{37} challenges~\cite{unix50} solvable by means of composing \unix pipelines.
By scanning GitHub, we were able to find a set of shell programs that solve all-but-one problems~\cite{}.
We consider this a good set of benchmarks because 
(i) the problems were designed to to highlight the \unix philosophy~\cite{} (rather than having a single command that solves the problem) and make extensive use of \unix built-ins with a plethora of flags, and (ii) the solutions were written by non-experts (contrary to the \S\ref{ours}) and were executed by \sys unmodified---stressing patterns that we, \sys's authors, did not necessarily anticipate.
In cases where an obvious fix improves the pipeline's performance, we report on both \sys's original speedup and the one achieved by applying the fix.
% TODO: talk about correct wrt developer intentions vs. correct wrt problem definition

% The ability to compose larger programs in versatile ways from smaller utility programs was a hallmark of the \unix design, and the game's levels are designed to stress that.
% These challenges were specifically designed to highlight the \unix's philosophy, its 
% ability to compose larger programs in versatile ways from smaller utility programs summarized by McIllroy's Unix philosophy memo.

@@

@@

\heading{Parallel Sort}
We compare a single \todo{GNU} \ttt{sort} optimized by \sys with the
same \ttt{sort} with the \ttt{--parallel} flag set. The results are
shown in \Cref{fig:sort-parallel-comparison}. There are a couple
important points to note. First of all, \sys without the \ttt{eager}
optimization enabled performs comparably with \ttt{sort
  --parallel}. On the other hand, enabling the \ttt{eager}
optimization allows \sys to perform much better (almost \todo{2}
times) than \ttt{sort} with the \ttt{--parallel} flag. That is because
\sys with the \ttt{eager} optimization enabled essentially uses
intermediate buffers (on disk) between the merge phases. On the other
hand, \ttt{sort --parallel} sorts the input trying to minimize
additional memory (so doesn't allocate intermediate buffers). \kk{If
  we want, we can add here numbers where we allow sort --parallel to
  use additional ram for intermediate buffers (using the -S flag).} Of
course, it would be possible to manually create a longer pipeline and
use intermediate buffers to speed this up, but this defeats the
purpose of having a simple solution.

@@

\heading{GNU Parallel}
We compare \sys to \ttt{parallel} (v.20160422), a GNU utility for running other commands in parallel~\cite{Tange2011a}.
In the first experiment, we leverage \ttt{parallel}'s \ttt{--jobs} flag and redirect input from a file to 
Sequential execution takes \todo{XX}s, and \sys's parallel execution takes \todo{XX}s (speedup: \todo{XX}\%).
We note that this pipeline is harsh for \sys, in that most of the overhead comes from a single command---masking \sys's improvements in other parts of the pipeline.

@@

There are a few possible ways one might attempt to use GNU \ttt{parallel} to parallelize this program.
Assuming a user knows that the bottleneck comes from the \ttt{XX} stage, they could attempt to use \ttt{parallel} there.
This leads to a runtime of \todo{XX}s (speedup: \todo{XX}\%).

Alternatively, a user could (incorrectly) sprinkle \ttt{parallel} across the entire program;
  this is a common strategy simplified by \ttt{parallel}'s \ttt{stdin} redirection feature.
Doing this leads to \todo{XXX}\% performance improvements but severely incorrect results with respect to the sequential execution---a \ttt{diff} that averages about \todo{XX--XX} lines.
\sys's conservative program transformations will not attempt parallelizing program sections that contain commands with unclear \todo{parallelizability} properties.
	% or commands whose inputs depend on earlier outputs.
GNU \ttt{parallel} is oblivious to command semantics, risking breaking program semantics as in this case.

% Parallel is a great program for its target audience.

\heading{Manual Parallelization}
Alternatively, a user might attempt to manually parallelize individual commands by using the abstractions provided by the POSIX shell.
The benefits depend on several factors, such as whether a command is I/O- or CPU-bound.

@@

\heading{Take-away}
\sys does not improve performance when running the risk of breaking correctness. % This should go in the intro
% Compared to \sys, \ttt{parallel} can lead to a combination of smaller performance improvements, incorrect execution with respect to the sequential program, and higher rewriting effort.
In highly-parallel benchmarks, \sys can lead to performance that is comparable to highly specialized streaming frameworks---at a fraction of the cost in terms of manual development.
